Making Sequential Programs Deterministic

### 2019-04-06 randoop-case-study.tex
Describe false positives

\item[Moderate issues]
  \strut
\begin{description}
\item[\bugHashMapOutput]
While printing diagnostic output, Randoop iterated over a \<HashMap> in
arbitrary order, making the output difficult to compare across different
executions, or even to find information in.
The class already implemented \<Comparable>, so
the developers changed \<methodWeights.keySet()> to
\codeid{new TreeSet<>(methodWeights.keySet()} in a \<for> loop (commit
\todo{TO DO}).
% This was actually introduced after the coding sprint.  Maybe some other
% defects were too, so our criticism should be about regressions rather
% than not finding them in the first place.
\end{description}

@@

\item[\bugHashcodeOutput]
  Diagnostic output printed a hash code for brevity.
  The developers changed it to have deterministic output (commit 661a4970).
  Although the Randoop developers classified this issue as minor, it is
  very similar to ones they fixed during their two-week coding sprint.

@@

\todo{Java 8 lambdas could help.}

>>>>

\item[Moderate issues]
  \strut
\begin{description}
\item[\bugHashMapOutput]
While printing diagnostic output, Randoop iterated over a \<HashMap> in
arbitrary order, making the output difficult to compare across different
executions, or even to find information in.
The class already implemented \<Comparable>, so
the developers changed \<methodWeights.keySet()> to
\codeid{new TreeSet<>(methodWeights.keySet())} in a \<for> loop (commit
f212cc7e).
% This was actually introduced after the coding sprint.  Maybe some other
% defects were too, so our criticism should be about regressions rather
% than not finding them in the first place.
\end{description}

@@

\item[\bugHashcodeOutput]
  Diagnostic output printed a hash code for brevity.
  The developers changed it to have deterministic output (commit 661a4970).
  Although the Randoop developers classified this issue as minor, it is
  similar to ones they fixed during their two-week coding sprint.

@@

\TheDeterminismChecker issued 37 false positive warnings, or 1 every 650
lines of code.

The most common reasons (responsible for 67\% of false positive warnings)
were:

\begin{itemize}
\item   (22\%)
%      8 constructors guarantee all instances of this class are @Det
  When every instance of a class is \<@Det>, \<@PolyDet> is equivalent to
  \<@Det>, but \theDeterminismChecker does not yet recognize this fact.
\item  (16\%)
%      6 https://github.com/t-rasmud/checker-framework/issues/96
  For a type that cannot be \<@OrderNonDet>, \<@PolyDet("up")> is
  equivalent to \<@PolyDet>.
\item (14\%)
%      5 can't specify receiver for anonymous class
  Java does not have syntax for specifying the type of an anonymous class.
\item (11\%)
%      4 actual type toString returns @PolyDet
  A variable is declared as \<Object> which has nondeterministic \<toString>,
  but at run time the value always has deterministic \<toString>.  For
  example, the run-time value is always an \<Integer> or a \<String>.
\item (5\%)
%      2 iterating over @OrderNonDet collection to modify another
  Iteration over an \<@OrderNonDet> collection in order to make another
  \<@OrderNonDet> collection, but the extracted elements have type
  \<@NonDet>.  For example, the following code is safe,
  \theDeterminismChecker issues a warning.  Variable \<elt> has type
  \<@NonDet>, so the call to \<add> does not typecheck.
\begin{Verbatim}
  void m(@OrderNonDet List<@Det String> input) {
    @OrderNonDet List<@Det String> output = new List<>();
    for (String elt : input) {
      output.add(elt);
    }
  }
\end{Verbatim}

\end{itemize}

The first two issues can be resolved by improvements to
\theDeterminismCheckerImplementation.
The third and fourth issues are limitations of the Java language, but by a
local refactoring in the Randoop codebase can make the code typecheck.
Refactoring can also eliminate the last warning.
